### 2018-07-25 sections/title_authors.tex

%--------------------------------------------------------%
%	TITLE SECTION with authors
%--------------------------------------------------------%

% Article title
  \title{\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}}

% Authors and Affiliations
  \author[1]{\large First Author}
  \author[1]{\large Second Author}
  \author[2]{\large Third Author}
  \affil[1]{\normalsize Department of XYZ, The University of ABC}
  \affil[2]{\normalsize Department of ABC, The University of XYZ}
  \renewcommand\Authands{ and }

% Today's date
	\date{Submitted: \usvardate\today}