### 2018-07-25 sections/title_noauthors.tex

%-----------------------------------------------------------
%	TITLE SECTION - with authors
%-----------------------------------------------------------

% Article title
  \title{\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}}

% Today's date
	\date{Submitted: \usvardate\today}

\makeatletter
\renewcommand{\@maketitle}{
\newpage
 \null
 \vskip 2em%
 \begin{center}%
  {\LARGE \@title \par}%
 \end{center}%
 \par} \makeatother