### 2018-08-16 sections/coverpage.tex
Update title options

%!TEX root = ../main.tex

\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

  \center % Center everything on the page

% HEADING SECTION

  %\textsc{\LARGE University Name}\\[1.5cm] % Name of your university/college
  % \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
  % \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title

% TITLE SECTION

  \vspace*{\fill}
  { \huge  1. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  { \huge  2. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  { \huge  3. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document

% AUTHOR SECTION

  % \vspace{1.5 cm}
  % Dileep Kishore\textsuperscript{1},
  % Gabriel Birzu\textsuperscript{2},
  % Zhenjun Hu\textsuperscript{1},
  % Charles DeLisi\textsuperscript{1,3},
  % Kirill Korolev\textsuperscript{1,2},
  % Daniel Segr\`{e}\textsuperscript{1,3,4}\\
  % \vspace{1cm}
  % \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
  % \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
  % \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
  % \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\

% % DATE SECTION

  % \vspace{1.5 cm}
  % {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise

  \vspace*{\fill} % Fill the rest of the page with whitespace

\end{titlepage}

%-----------------------------------------------------------------

\newpage