### 2019-11-25 sections/acknowledgments.tex
Remove supplementary for now

%!TEX root = ../main.tex

\section*{Acknowledgments}

\section*{Contributions}