### 2020-05-29 sections/acronyms.tex

%!TEX root = ../main.tex

\begin{acronym}[XXXXXXXX]
    \acro{ngs}[NGS]{next-generation sequencing}
    \acro{mind}[MIND]{Microbial Interaction Network Database}
    \acro{micone}[MiCoNE]{Microbial Co-occurrence Network Explorer}
    \acro{otu}[OTU]{Operational Taxanomic Unit}
    \acro{esv}[ESV]{Exact Sequence Variant}
    \acro{pcr}[PCR]{Polymerase Chain Reaction}
    \acro{gg}[GG]{Greengenes}
    \acro{ncbi}[NCBI]{National Center for Biotechnology Information}
    \acro{comets}[COMETS]{Computation of Microbial Ecosystems in Time and Space}
    \acro{dada2}[DADA2]{Divisive Amplicon Denoising Algorithm 2}
    \acro{qiime1}[QIIME1]{Quantitative Insights Into Microbial Ecology 1}
    \acro{qiime2}[QIIME2]{Quantitative Insights Into Microbial Ecology 2}
    \acro{sparcc}[SparCC]{Sparse Correlations for Compositional data}
    \acro{spieceasi}[SpiecEasi]{Sparse InversE Covariance estimation for Ecological Association and Statistical Inference}
    \acro{mldm}[mLDM]{metagenomic Lognormal-Dirichlet-Multinomial}
    \acro{magma}[MAGMA]{Microbial Association Graphical Model Analysis}
    \acro{kegg}[KEGG]{Kyoto Encyclopedia of Genes and Genomes}
    \acro{fba}[FBA]{Flux Balance Analysis}
    \acro{dc}[DC]{Denoising and Clustering}
    \acro{ta}[TA]{Taxonomy Assignment}
    \acro{op}[OP]{OTU Processing}
    \acro{ni}[NI]{Network Inference}
\end{acronym}