### 2020-08-25 sections/acknowledgments.tex
Updates from Overleaf

%!TEX root = ../main.tex

% TODO: Update this section
\section*{Acknowledgments}

We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript. This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S. Department of Energy, Office of Science, Office of Biological & Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.

\section*{Contributions}
Designed the research project: DK, KK, DS, ZH, CDL. Performed analysis: DK, GB. Wrote the first draft of the manuscript: DK. Revised and wrote final version of the manuscript: DK, DS, KK.