### 2020-08-25 sections/coverpage.tex

%!TEX root = ../main.tex

\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

  \center % Center everything on the page

% HEADING SECTION

  %\textsc{\LARGE University Name}\\[1.5cm] % Name of your university/college
  % \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
  % \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title

% TITLE SECTION

  \vspace*{\fill}
  {\huge Inferring microbial co-occurrence networks from amplicon data: a systematic evaluation}\\[0.4cm]
  % {\huge  2. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  % {\huge  3. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  % {\huge  4. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document

% AUTHOR SECTION

  \vspace{1.5 cm}
  Dileep Kishore\textsuperscript{1,2},
  Gabriel Birzu\textsuperscript{3},
  Zhenjun Hu\textsuperscript{1},
  Charles DeLisi\textsuperscript{1,3},
  Kirill Korolev\textsuperscript{$\dagger$1,2,3},\\
  Daniel Segr\`{e}\textsuperscript{$\dagger$1,2,4,5}\\
  \vspace{1cm}
  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{2}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{3}Department of Physics, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{4}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{5}Department of Biology, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

% % DATE SECTION

  % \vspace{1.5 cm}
  % {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise

  \vspace*{\fill} % Fill the rest of the page with whitespace

\end{titlepage}

%-----------------------------------------------------------------

\newpage