### 2022-09-08 sections/discussion.tex
Update on Overleaf.

%!TEX root = ../main.tex

\section*{Discussion}

  \subsection*{Why \ac{micone}?}

  A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
  Our analysis shows that networks generated using different combinations of tools and approaches can be significantly different from each other, highlighting the need for a clear assessment of the source of variability and for tools that provide the most robust and accurate results.
  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to compare networks generated by multiple possible combinations of tools and parameters.
  Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
  The \ac{micone} Python package provides functions and methods to perform a detailed analysis of the count matrices and the co-occurrence networks.
  The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.

  While several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} can be used to generate co-occurrence networks from 16S sequencing data, no single tool exist that integrates the complete process of inferring microbial interaction networks from 16S sequencing reads.
  \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster, or in the cloud.

  \subsection*{The default pipeline and recommended tools}

  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
  Additionally, we develop a consensus approach, that can reliably generate networks that are robust across multiple tool/parameter choices.
  An important caveat related to these results is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
  While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.

  The networks generated by different network inference methods show considerable differences in edge-density and connectivity, partially due to the underlying assumptions regarding sparsity, distribution and compositionality.
  To address this issue, we have developed two consensus algorithms (simple voting and scaled-sum method) that generate networks whose links have evidence based on multiple inference algorithms.

 We find that the scaled-sum method (along with p-value merging for correlation-based methods) performs the best on synthetic datasets, and is therefore chosen as default for the \ac{ni} step of the pipeline. Notably, the consensus network displays a higher number of true positives, while also decreasing the overall variability across the whole pipeline (Figure~\ref{fig:figure7}B). The concise list of robust associations revealed by the consensus algorithm represents a valuable set for experimental validation follow-up.

  \subsection*{Future directions}

  Future work building upon our current results could enhance the network inference process in multiple ways.
  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.

  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (e.g. Family).
  As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogenetically relevant comparisons.

  Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
  Further, benchmarking of co-occurrence networks could also be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
  Additionally, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.

  Although in the current analysis, we only use default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
  Overall, it is important to keep in mind that there likely is no ``best method'' for the various steps of 16S data analysis.
  Therefore, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.

  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
  It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and will require more systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.