### 2022-09-27 tables/figure_s8_caption.tex
Fix: Thresholds and alpha in Figure 4

\caption{
    \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data when Pearson and Spearman are included in the consensus calculation}.
    The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
    The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
    With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure_s6}.
    The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to an increase in the parameter value.
    The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
}